An Efficient Feedback Calibration Algorithm for Direct Imaging Radio Telescopes

### 2016-05-03 cal_demo/ms.tex
More smoothing of text, expanding on explanation of gridding step

Thus far we have assumed the antennas perfectly measure the sky electric field distribution according to their far-field radiation patterns. Here we introduce a multiplicative complex gain as well as an additive noise term which corrupt the measured signals.
\begin{equation}\label{eq:apply_gain}
E_a(f,t) = g_a(f,t) E_a^I(f,t) + n_a(f,t)
\end{equation}
The noise term is 
strictly receiver noise -- noise introduced by the instrument. Any sky noise is implicitly included 
in the time dependence of the sky electric field. 
We have carried the frequency dependence to this point to make clear the complex gains are
frequency dependent. However, each subsequent step treats each frequency 
channel independently, so we will drop the $f$ to simplify notation. 
Furthermore, while antenna gains may vary with time, we will assume they are constant over
the period of our calibration, and thus drop their time dependence from the notation.

The MOFF algorithm next calls for a calibration of the electric field measurements. The goal of our new calibration method will be 
to form an estimate of the gains, $g'_a$, where the tick represents the estimate. For now we will assume we have formed an estimate 
to proceed with the MOFF pipeline. We correct the incoming electric field data stream using our 
current estimate of the gains.
\begin{equation}
E'_a(t) = E_a(t)/ g'_a
\end{equation}

The next step is to grid the calibrated electric field measurements. MOFF uses the antenna patterns as the gridding 
kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
Mathematically we can express the gridding operation as
\begin{equation}
E'_j=\sum_a W_a(\mathbf{r}_i-\mathbf{r}_a) E'_a(t)
\end{equation}
This 
achieves an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
traditional correlators which grid visibilities, the MOFF grids the electric fields directly. This 
operation must be performed for every time step, before any averaging of the data.

>>>>

Thus far we have assumed the antennas perfectly measure the sky electric field distribution according to their far-field radiation patterns. Here we introduce a multiplicative complex gain as well as an additive noise term which corrupt the measured signals.
\begin{equation}\label{eq:apply_gain}
E_a(f,t) = g_a(f,t) E_a^I(f,t) + n_a(f,t)
\end{equation}
%The noise term is 
%strictly receiver noise -- noise introduced by the instrument. Any sky noise is implicitly included 
%in the time dependence of the sky electric field. 
We have carried the frequency dependence to this point to make clear the complex gains are
frequency dependent. However, each subsequent step treats each frequency 
channel independently, so we will drop the $f$ to simplify notation. 
Furthermore, while antenna gains may vary with time, we will assume they are constant over
the period of our calibration, and thus drop their time dependence from the notation.

The MOFF algorithm next calls for a calibration of the electric field measurements. The goal of our new calibration method will be 
to form an estimate of the gains, $g'_a$, where the prime represents an estimate. For now we will assume we have formed an estimate 
to proceed with the MOFF pipeline. We correct the incoming electric field data stream using our 
current estimate of the gains.
\begin{equation}
E'_a(t) = E_a(t)/ g'_a
\end{equation}

The next step is to grid the calibrated electric field measurements to regularly spaced 
positions on the 2D plane of the ground, $\mathbf{r}_j$. 
MOFF uses the antenna patterns as the gridding 
kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
Mathematically we can express the gridding operation as
\begin{equation}
E'_j=\sum_a W_a(\mathbf{r}_j-\mathbf{r}_a) E'_a(t)
\end{equation}
This 
achieves an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
traditional correlators which grid visibilities, the MOFF grids the electric fields directly. This 
operation must be performed for every time step, before any averaging of the data.
